---
title: "Summary"
author: "JJayes"
date: '2022-04-19'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose

A summary of the electrification committee reports produced between 1924 and 1926 by the electrification committee, focussing on what data to extract.

## Which regions?

```{r}
c("Blekinge",
  "Jämtland.pdf",
  "Södermanland.pdf",
  "Västerbotten.pdf",
  "Värmland.pdf",
  "Gotland.pdf",
  "Jönköping.pdf",
  "Kristianstad.pdf",
  "Kronobergs",
  "Örebro.pdf",
  "Koppaberg.pdf",
  "Norrbottens",
  "Västernorrland.pdf",
  "Halland.pdf",
  "Uppsala.pdf",
  "Skaraborgs",
  "Kalmar",
  "Östergötland.pdf",
  "Älvsborg.pdf",
  "Goteborgs och Bohus",
  "Västmanland.pdf",
  "Stockholm.pdf",
  "Gävleborg.pdf") %>% 
  as_tibble() %>% 
  mutate(value = str_remove(value, "\\.pdf"))
```


## What are the contents of each report??

- Introduction
- Description of current electrical installations
- Description of resources that remain to be tapped
- Analysis of the needs in the county and projection into the future
- Propposal for planned electrification
  - Sources of power
  - Local distribution
  - Costs of electrification
- Summary

A big table with very detailed information about the power generation in each little town! Wow!

### What tables do we want to get?

- Full consumption estimate and probable electrification

Missing Gotland.
Missing Kristianstad

- Water vs Heat power:

Missing Gotland, Halland,
