---
title: "Electrification summary"
format: html
---

## Purpose

EDA on electrification summary transcribed with textract.

```{r}
library(tidyverse)
library(here)
theme_set(theme_light())
```

## Map init

```{r}
library(sf)
map <- read_rds(here("data", "county_map.rds"))

map %>% 
  ggplot() +
  geom_sf() +
  geom_sf_label(aes(label = lan_name))
```


```{r}
library(ggiraph)

gg_iraph_map <- map %>% 
    ggplot() +
    geom_sf_interactive(aes(tooltip = lan_name)) +
    theme_void()

girafe(
  ggobj = gg_iraph_map,
  width_svg = 8,
  height_svg = 4,
  options = list(
    # opts_tooltip(css = tooltip_css, delay_mouseover = 0, delay_mouseout = 0),
    opts_hover_inv(css = "opacity:0.1;"),
    opts_hover(css = "stroke-width:2;")
  )
)
```

### Read in data

Data is in two tables with two sets of codebooks.

Starting with table 1, about costs and such.

```{r}
df <- readxl::read_excel(here("data/electricity_summary_table_1.xlsx"))

df <- df %>% 
  janitor::clean_names() %>% 
  pivot_longer(-c(x1, x2)) %>% 
  mutate(name = parse_number(name)) %>% 
  rename(index = x1, county = x2, col = name)

df <- df %>% 
  filter(!county %in% c("Summa", "check"),
         !is.na(county))

codebook <- readxl::read_excel(here("data/Electricity_summary_table_1_codebook.xlsx"))

df <- df %>% 
  inner_join(codebook, by = "col")

```

```{r}
#| column: page
df %>% 
  select(county, desc, value) %>% 
  pivot_wider(names_from = desc, values_from = value) %>% 
  gt::gt()
```

### Sections

```{r}
df %>% distinct(desc) %>% pull()
```

Geographic stuff:
- Total land area A1. 
- Field area A0
- Population and pop density* check

Farm features
- Average farm size
- Farm size vs pop density.



### Geographic area

```{r}
df %>%
  filter(desc %in% c("Total land area A1", "Field area A0", "Forest area As")) %>% 
  mutate(county = fct_reorder(county, value)) %>% 
  ggplot(aes(value, county, fill = desc)) +
  geom_col(position = "dodge", show.legend = F) +
  scale_x_continuous(labels = scales::number_format()) +
  facet_wrap(~ desc, scales = "free")
```

I should do a shiny app with four panels, one being a table with coloured columns...

### Electrified field area

```{r}
df %>%
  filter(str_detect(desc, "Electrified field area: as a")) %>%
  mutate(county = fct_reorder(county, value)) %>%
  ggplot(aes(value, county, fill = desc)) +
  geom_col(position = "dodge", show.legend = F) +
  scale_x_continuous(labels = scales::percent_format(scale = 1)) +
  facet_wrap(~desc, scales = "free")

gg_iraph_map <- df %>%
  filter(str_detect(desc, "Electrified field area: as a")) %>% 
  inner_join(map, by = c("county" = "lan_name")) %>%
  mutate(tooltip = str_c(county, "\n", value, " %")) %>% 
  ggplot() +
  geom_sf_interactive(aes(tooltip = tooltip, data_id = county,
                          fill = value, geometry = geometry)) +
  scale_fill_gradient2(
    low = "blue",
    high = "red",
    mid = "pink",
    midpoint = 40
  ) +
  theme_void() +
  labs(fill = "Electrified field area %")

girafe(
  ggobj = gg_iraph_map,
  width_svg = 8,
  height_svg = 4,
  options = list(
    # opts_tooltip(css = tooltip_css, delay_mouseover = 0, delay_mouseout = 0),
    opts_hover_inv(css = "opacity:0.1;"),
    opts_hover(css = "stroke-width:2;")
  )
)
```

For some regions we can - distinguishing between lighting and motor:

```{r}
counties <- df %>%
  filter(str_detect(desc, "Electrified field area: with"),
         !is.na(value)) %>% 
  distinct(county)

df_electvar <- df %>%
  filter(
    county %in% counties$county,
    str_detect(desc, "Electrified field area: with|Field area A0")
  ) %>%
  select(county, value, desc) %>%
  pivot_wider(names_from = desc, values_from = value) %>%
  mutate(across(starts_with("Elect"), ~ .x / `Field area A0`)) %>%
  mutate(county = 
           fct_reorder(county, `Electrified field area: with lighting and motors`)) %>%
  pivot_longer(-c(county, `Field area A0`)) %>%
  mutate(name = fct_relevel(name, "Electrified field area: with lighting only"))

# df_electvar %>% write_rds("data/df_electvar.rds")

df_electvar %>% 
  ggplot(aes(value, county, fill = name)) +
  geom_col(position = "stack") +
  scale_x_continuous(labels = scales::percent_format(scale = 100)) +
  labs(x = "Percentage of arable land")
```


### Rural consumption

```{r}
#| column: page

df %>%
  filter(str_detect(desc, "Rural consumption:")) %>%
  mutate(
    desc = str_remove(desc, "Rural consumption: "),
    desc = str_c(desc, " - ", units),
    desc = str_to_title(desc)
  ) %>%
  select(county, desc, value) %>%
  pivot_wider(names_from = desc, values_from = value) %>%
  arrange(desc(`Electrification Degree Currently - Pct`)) %>%
  gt::gt()
```



Cost per hectare for rural electrification: Electrification plant performed so far: Costs reduced to the 1914 price situation

```{r}

gg_iraph_map_cost <- df %>%
  filter(str_detect(desc, "Cost per hectare for rural electrification: Electrification plant performed so far: Costs reduced to the 1914 price situation")) %>%
  inner_join(map, by = c("county" = "lan_name")) %>%
  mutate(tooltip = str_c(county, "\n", value, " kronor")) %>% 
  ggplot() +
  geom_sf_interactive(aes(tooltip = tooltip, 
                          data_id = county,
                          fill = value, geometry = geometry)) +
  scale_fill_gradient2(
    low = "purple",
    high = "orange",
    mid = "yellow",
    midpoint = 100,
    labels = scales::dollar_format(suffix = "Kr", prefix = "")
  ) +
  theme_void() +
  labs(fill = "Cost per hectare for rural electrification")

girafe(
  ggobj = gg_iraph_map_cost,
  width_svg = 8,
  height_svg = 4,
  options = list(
    # opts_tooltip(css = tooltip_css, delay_mouseover = 0, delay_mouseout = 0),
    opts_hover_inv(css = "opacity:0.1;"),
    opts_hover(css = "stroke-width:2;")
  )
)
```

Combined

```{r}
library(cowplot)

int_fig <- girafe(
  ggobj = plot_grid(gg_iraph_map, gg_iraph_map_cost, nrow = 2),
  width_svg = 5,
  height_svg = 3,
  options = list(
    # opts_tooltip(css = tooltip_css, delay_mouseover = 0, delay_mouseout = 0),
    opts_hover(css = "fill: #E69F00; fill-opacity: 1.0; stroke: #E69F00;")
  )
)

# write_rds(int_fig, "data/interactive_fig.rds")

```



Cost vs 

```{r}
df %>%
  filter(str_detect(desc, "Electrified field area: as a|Cost per hectare for rural electrification: Electrification plant performed so far: Costs reduced to the 1914 price situation")) %>%
  select(county, value, desc) %>%
  pivot_wider(names_from = "desc") %>%
  ggplot(aes(
    `Electrified field area: as a percentage of total arable land area`,
    `Cost per hectare for rural electrification: Electrification plant performed so far: Costs reduced to the 1914 price situation`
  )) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_label(aes(label = county)) +
  scale_x_continuous(labels = scales::percent_format(scale = 1)) +
  scale_y_continuous(labels = scales::dollar_format(prefix = "", suffix = " Kr")) +
  labs(y = "Cost per hectare for rural electrification (1914 prices)")
```


## Table 2

```{r}
df <- readxl::read_excel(here("data/electricity_summary_table_2.xlsx"))

df <- df %>% 
  janitor::clean_names() %>% 
  pivot_longer(-c(x1, x2)) %>% 
  mutate(name = parse_number(name)) %>% 
  rename(index = x1, county = x2, col = name)

df <- df %>% 
  filter(!county %in% c("Summa för hela landet", "check"),
         !is.na(county))

codebook <- readxl::read_excel(here("data/Electricity_summary_table_2_codebook.xlsx"))

df <- df %>% 
  inner_join(codebook, by = "col")
```


```{r}
#| column: page

df %>% 
  select(county, desc, value) %>%
  pivot_wider(names_from = desc, values_from = value) %>% 
  gt::gt()
```

```{r}
df %>% 
  filter(desc %in% c("Number of power stations", "Total power")) %>% 
  mutate(county = fct_reorder(county, value)) %>% 
  ggplot(aes(value, county)) +
  geom_col() +
  facet_wrap(~ desc, scales = "free")
```
