---
title: "Grid Summary"
format: html
---

## Purpose

Summary of the expansion of the electricity grid in Sweden in the years 1885-1926.

### Maps of the grid and production by type

If we look at the production in 1885 - it looks like this:

```{r}
library(tidyverse)
library(sf)

st_map <- read_rds(here::here("data", "maps", "1900_parish_map.rds"))

production_1885 <- read_rds(here::here("data", "maps", "1885_water_and_heat.rds"))

ggplot() +
  geom_sf(data = st_map, fill = "lightgrey", colour = "grey80", alpha = .5) +
  geom_sf(data = production_1885, aes(colour = type, size = power)) +
  theme_minimal() +
  labs(size = "Power in kWh",
       colour = "")

```

The in 1900 we get the production and the grid:

```{r}
production_1900 <- read_rds(here::here("data", "maps", "1900_water_and_heat.rds"))

grid_1900 <- read_rds(here::here("data", "maps", "1900_grid.rds"))

ggplot() +
  geom_sf(data = st_map, fill = "lightgrey", colour = "grey80", alpha = .5) +
  geom_sf(data = production_1900, aes(colour = type, size = power)) +
  geom_sf(data = grid_1900, colour = "red") +
  theme_minimal()  +
  labs(size = "Power in kWh",
       colour = "")
```

Then in 1911 we can overlap the 1911 grid with the 1900 production and see that the grid has expanded in the mean time.

```{r}
grid_1911 <- read_rds(here::here("data", "maps", "1911_grid.rds"))

ggplot() +
  geom_sf(data = st_map, fill = "lightgrey", colour = "grey80", alpha = .5) +
  geom_sf(data = production_1900, aes(colour = type, size = power)) +
  geom_sf(data = grid_1911, colour = "red") +
  theme_minimal() +
  labs(size = "Power in kWh",
       colour = "")
```

What about the grid in 1926?

```{r}
grid_1926 <- read_rds(here::here("data", "maps", "1926_grid.rds"))

ggplot() +
  geom_sf(data = st_map, fill = "lightgrey", colour = "grey80", alpha = .5) +
  # geom_sf(data = production_1900, aes(colour = type, size = power)) +
  geom_sf(data = grid_1926, colour = "red") +
  theme_minimal()
```

Now we need to get the information about the power stations in 1926 from the different electrification committee reports about rural electrification.

